---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
data <- read.table(unz("activity.zip", "activity.csv"), header=TRUE, sep=",")
data <- data[complete.cases(data), ]
```

## What is mean total number of steps taken per day?

To find out we calculate the total number of steps taken per day and draw a histogram showing the frequency distribution.

```{r}
total_steps_per_day <- aggregate(steps ~ date, data = data, FUN = sum)
hist(total_steps_per_day$steps, main = "Histogram: Total number of steps taken each day", xlab = "Total number of steps per day")
```

Now we calculate and show the mean and median of the total number of steps taken per day:

```{r}
print(paste("Mean of total steps per day:", mean(total_steps_per_day$steps)))
```
```{r}
print(paste("Median of total steps per day:", median(total_steps_per_day$steps)))
```


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
